---
title: "16S"
author: "Nicholas Baetge"
date: "8/13/2020"
output: github_document
---

# Intro

Here, the NAAMES cast 16S sequences processed by Luis Bolanos in the Giovannoni group are explored

```{r message = F, warning = F}
library(tidyverse) 
library(rmarkdown)
library(knitr)
library(readxl)
library(data.table) 
library(scales)
library(zoo)
library(oce)
library(patchwork)
#rmarkdown tables
library(stargazer)
library(pander)
#stat tests
library(lmtest)
library(lmodel2)
library(rstatix)
library(ggpubr)
#for odv type plots
library(lubridate)
library(reshape2)
library(MBA)
library(mgcv)
#phyloseq
library(phyloseq)

custom_theme <- function() {
  theme_test(base_size = 30) %+replace%
    theme(legend.position = "right",
          legend.spacing.x = unit(0.5,"cm"),
          legend.title = element_text(size = 14),
          legend.text = element_text(size = 14),
          legend.background = element_rect(fill = "transparent",colour = NA),
          legend.key = element_rect(fill = "transparent",colour = NA),
          panel.background = element_rect(fill = "transparent",colour = NA),
          plot.background = element_rect(fill = "transparent",colour = NA)) 
}

custom.colors <- c("AT39" = "#377EB8", "AT34" = "#4DAF4A", "AT38" = "#E41A1C", "AT32" = "#FF7F00", "Temperate" = "#A6CEE3", "Subpolar" = "#377EB8", "Subtropical" = "#FB9A99", "GS/Sargasso" = "#E41A1C", "Early Spring" = "#377EB8", "Late Spring" = "#4DAF4A","Early Autumn" = "#E41A1C", "Summer" = "#E41A1C", "Late Autumn" = "#FF7F00", "Gv2_2019" = "#377EB8", "WOA18_MN" = "#4DAF4A", "WOA18_AN" = "#E41A1C")

levels = c("GS/Sargasso", "Subtropical", "Temperate", "Subpolar",  "AT39-6", "AT34", "AT38", "AT32","South", "North", "Early Spring", "Late Spring","Early Autumn",  "Summer", "Late Autumn", "Gv2_2019", "WOA18_MN", "WOA18_AN","Nov", "Nov sd", "Dec", "Dec sd", "Jan", "Jan sd", "Feb", "Feb sd", "Mar", "Mar sd", "Apr", "Apr sd",  "Cruise", "ARGO", "5-75 m", "100-200 m", "300 m")


bar.colors <- c("100 m" = "white", "CM" = "#4DAF4A",  "PAM" = "#377EB8")

odv.colors <- c("#feb483", "#d31f2a", "#ffc000", "#27ab19", "#0db5e6", "#7139fe", "#d16cfa")
```

# Import Data 

```{r message = F}

count.tab <- read.table("~/GITHUB/naames_multiday/Input/16s/HetV1OTU.txt", header = T, row.names = 1, check.names = F)

tax.tab <- as.matrix(read.table("~/GITHUB/naames_multiday/Input/16s/HetV1TUtax.txt", header = T, row.names = 1, check.names = F, na.strings = "", sep = "\t"))

sample.tab <- read_rds("~/GITHUB/naames_multiday/Input/export_ms/processed_bf.2.2020.rds") %>% 
  select(Cruise:CampCN, Target_Z, DNA_ID) %>% 
  drop_na(DNA_ID) %>% 
  rename(z = Target_Z) %>% 
  left_join(., read_rds("~/GITHUB/naames_multiday/Output/processed_data.rds") %>%
              select(Cruise, Station, Date,  CampCN, mld, z, chl, doc, o2, n, phyc, bc, bcd ) %>% 
              distinct() %>% 
              mutate_at(vars(phyc:bcd), function(x)(x/10^3))) %>% 
  mutate(`Depth Interval` = ifelse(z <= 75, "5-75 m", "100-200 m"),
         `Depth Interval` = ifelse(z == 300, "300 m", `Depth Interval`)) %>% 
  select(Cruise:z, `Depth Interval`, everything()) %>% 
  column_to_rownames(var = "DNA_ID")

```

# Phyloseq Object

We need to create a phyloseq object that merges all three datasets. Sometimes this doesn't work beacuse of the format of the data files. Make sure all the sample names between the sampleinfo.txt and seqtab-nochimtaxa.txt are the same

```{r}
OTU = otu_table(count.tab, taxa_are_rows = TRUE) 
TAX = tax_table(tax.tab)
SAM = sample_data(sample.tab)
ps = phyloseq(OTU,TAX,SAM) 

sample_data(ps)$`Depth.Interval` <- factor(sample_data(ps)$`Depth.Interval`, levels = levels)
sample_data(ps)$Season <- factor(sample_data(ps)$Season, levels = levels)
```

We can rarefy our data at this point, though there is debate over this. 

```{r}
rarefy_ps = rarefy_even_depth(ps, sample.size = min(sample_sums(ps)), rngseed = 42398403)
```


# Custom Table

phyloseq objects can sometimes be hard to handle when you have a particular plot in mind that you want to make to make it easier for us: we'll extract the relative abundance data from the object and merge them with the taxa and sample info data as a new dataframe


## Generate relative abundances

Our data currently shows the relative proportion of different sequences to the total number of gene copies recovered, so we'll normalize the gene copy number

```{r}
ps_std <- transform_sample_counts(ps, function(x) x/sum(x))
#extract the relative abundance table and coerce into dataframe
ps_std.tab <- as(otu_table(ps_std), "matrix")
ps_std.df = as.data.frame(ps_std.tab) 
```

## Table with Rel Abund, Taxa, Sample Info 

Create a new table that combines relative abundances with the taxa table

```{r}
#first coerce the taxa table into a data frame
tax.df = as.data.frame(tax.tab) 
#then combine the data frames
custom.tab <- tax.df %>% 
  rownames_to_column(., var = "asv") %>% 
  left_join(., ps_std.df %>% rownames_to_column(., var = "asv")) %>% 
  #create a new index of that combines the  class, order, family, and genus values
  mutate(cofg = paste(Class, "_", Order,"_", Family, "_", Genus )) %>% 
  select(-c(asv:Genus)) %>% 
  select(cofg:everything()) 

#save the row names and then make them into the column names
colnames <- custom.tab[,1]

#transpose the dataframe so we can merge with the sample info table
t_custom.tab <-  as.data.frame(t(custom.tab[,-1]))
colnames(t_custom.tab) <- colnames

#merge
sweet.tab <- t_custom.tab %>% 
  rownames_to_column(., var = "sample") %>% 
  left_join(., sample.tab %>% rownames_to_column(., var = "sample")) %>% 
  select(sample, Cruise:bcd, everything()) %>% 
  arrange(CampCN, z)
```

## Save

```{r}
saveRDS(sweet.tab, "~/GITHUB/naames_multiday/Output/Custom_ASV_Table.rds")
```


# Ordinate Samples

```{r}
ord <- ordinate(ps, method = "PCoA", distance = "bray")
```

```{r echo = FALSE, warning = FALSE, message = FALSE, fig.height = 10, fig.width = 14, fig.align = "center"}

ord.plot <- plot_ordination(ps, ord,  title = "Bray Curtis PCoA", shape = "Depth.Interval", color = "Latitude") +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_point(alpha = 0.6, stroke = 2, size = 4) +
  geom_text(aes(label = Station), hjust = 0, vjust = 0, color = "black", alpha = 0.7, size = 3, position = position_jitter(), fontface = "bold") +
  scale_color_gradientn(colors = odv.colors) +
  custom_theme()  
#removing one of the plotting layers (there are points within points)
ord.plot$layers <- ord.plot$layers[-1]

ord.plot + facet_wrap(~Season) +
   guides(color = guide_colorbar(title = "Latitude, ˚N"), shape = guide_legend(title = "Depth Interval"))
```














