---
title: "Floats"
author: "Nicholas Baetge"
date: "12/23/2021"
output: github_document
---

# Intro

Analysis of [floats deployed](http://misclab.umeoce.maine.edu/floats/float/n0647) at N2S4-anticyclonic eddy (n0647) and N3S6-no eddy (n0847, n0846), at N3S3-mode water eddy (n0850), and at N3S4-cyclonic eddy (n0849). Also include ship data for N3S3.5-no eddy, which has no float data associated. 

- Input data, which were the level 2 processed data, were pulled from the [NAAMES data page](https://naames.larc.nasa.gov/data2018.html)
- MLDs were calculated using the N^2^ buoyancy frequency and are defined as the depth below 5 m at which N^2^ > |stdev(N^2^)|

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(lubridate)
library(hms)
library(zoo) 
library(oce)  
library(ggpubr)
library(patchwork)

#for mapping
library(akima)
library(mapdata)
library(maps)
library(rerddap)

#for T-S
library(PlotSvalbard)


```

# Import and Wrangle Data

```{r message = F}
#pass all the filenames of everything in a float folder into a variable
# filenames <- list.files("~/GITHUB/naames_multiday/Input/n0849/",pattern = "*.csv", full.names = T)
# profile.names <- list.files("~/GITHUB/naames_multiday/Input/n0849/",pattern = "*.csv", full.names = F)
# profile.number <- gsub('.csv', '', profile.names) %>%
#   gsub('n0849.', '', .) %>%
#   as.numeric(.)

#now go through all the files in the folder, pick out the filenames, and pass them into the variable 'filenames'. use the file names to interate through the files, extracting the data from each csv, and combining it into one list
# master <-  lapply(filenames, function(i){
#   read_csv(i)
# })
# names(master) <- profile.number

#convert to data frame
# profiles <- bind_rows(master, .id = "profile")

#save csv
# write_csv(profiles, "~/GITHUB/naames_multiday/Input/n0849_profiles.csv")
```


## drifter


```{r message=FALSE}
drifters <- read_csv("~/GITHUB/naames_multiday/Input/drifters/N2S4.csv") %>%
  mutate(Cruise = "AT34", name = "NAAMES 2", Station = 4, drifter = 1, plot_sep = "Subtropical") %>% 
  bind_rows(., read_csv("~/GITHUB/naames_multiday/Input/drifters/N3S6_01.csv") %>%
  mutate(Cruise = "AT38", name = "NAAMES 3", Station = 6, drifter = 1, plot_sep = "Subpolar")) %>% 
  bind_rows(., read_csv("~/GITHUB/naames_multiday/Input/drifters/N3S6_02.csv") %>%
  mutate(Cruise = "AT38", name = "NAAMES 3", Station = 6, drifter = 2, plot_sep = "Subpolar")) %>% 
  mutate(plot_name = paste(name, "Station", Station), 
         type = "Drifter") %>% 
  rename(lat = latitude_deg_,
         lon = longitude_deg_,
         Date = timestamp_utc_) %>% 
  select(Cruise, name, Station, plot_name, plot_sep, type, drifter, lat, lon, Date)
```

## floats

```{r warning = FALSE, message = FALSE}

n0647 <- read_csv("~/GITHUB/naames_multiday/Input/n0647_profiles.csv") %>%  
  mutate(float = "n0647", Cruise = "AT34", Station = 4) %>% select(float:Station, profile, datetime:s, par, o2_c, chla_adj, fchl, fdom, cphyto, poc) %>%
  rename(o2 = o2_c, chla = chla_adj) %>% 
  filter(datetime < ymd_hms("2016-06-03 06:35:00"))

n0846 <- read_csv("~/GITHUB/naames_multiday/Input/n0846_profiles.csv") %>%  
  mutate(float = "n0846", Cruise = "AT38", Station = 6) %>% select(float:Station, profile, datetime:s, par, o2_c, chla_adj, fchl, fdom, cphyto, poc) %>%
  rename(o2 = o2_c, chla = chla_adj) %>% 
  filter(datetime < ymd_hms("2017-09-21 00:00:00"))

n0847 <- read_csv("~/GITHUB/naames_multiday/Input/n0847_profiles.csv") %>%  
  mutate(float = "n0847", Cruise = "AT38", Station = 6) %>% select(float:Station, profile, datetime:s, par, o2_c, chla_adj, fchl, fdom, cphyto, poc) %>%
  rename(o2 = o2_c, chla = chla_adj) %>% 
  filter(datetime < ymd_hms("2017-09-21 00:00:00"))

n0850 <- read_csv("~/GITHUB/naames_multiday/Input/n0850_profiles.csv") %>%  
  mutate(float = "n0850", Cruise = "AT38", Station = 3) %>% select(float:Station, profile, datetime:s, par, o2_c, chla_adj, fchl, fdom, cphyto, poc) %>%
  rename(o2 = o2_c, chla = chla_adj) %>% 
  filter(datetime < ymd_hms("2017-09-08 20:00:00"))

n0849 <- read_csv("~/GITHUB/naames_multiday/Input/n0849_profiles.csv") %>%  
  mutate(float = "n0849", Cruise = "AT38", Station = 4) %>% select(float:Station, profile, datetime:s, par, o2_c, chla_adj, fchl, fdom, cphyto, poc) %>%
  rename(o2 = o2_c, chla = chla_adj) %>% 
  filter(datetime < ymd_hms("2017-09-11 01:00:00"))

```

# Calculations

## MLDs and EZD

Euphotic zone depth calculated as depth at which PAR is 1% surface PAR. 

```{r warning = FALSE, message = FALSE}

#combine all the float data
naames_floats <- rbind(n0647, n0846, n0847, n0850, n0849) 

floatsNmlds <-  naames_floats %>% 
  group_by(float, profile, datetime) %>% 
  mutate(sigT = oce::swSigmaTheta(salinity = s, temperature = t, pressure = p, referencePressure = 0),
         potT = swTheta(salinity = s, temperature = t, pressure = p),
         N2 = oce::swN2(pressure = p, sigmaTheta = sigT)) %>% 
  filter(p > 5 & N2 > abs(sd(N2))) %>% 
  filter(p == min(p)) %>% #report the shallowest depth at which the above condition is met
  rename(MLD = p) %>% 
  ungroup() %>% 
  select(float:MLD) %>% 
  distinct() %>% 
  left_join(naames_floats, .) 
  
  # filter(datetime == ymd_hms("2016-05-24 12:28:04")) %>%
  
floatsNmldsNezds <- floatsNmlds %>% 
  group_by(float, profile, datetime) %>% 
  mutate(onePercSurfPAR = 0.01 * first(par)) %>% 
  group_modify(~ add_row(., par = max(.$onePercSurfPAR))) %>% 
  fill(c(Cruise, Station, lat, lon, MLD, onePercSurfPAR)) %>% 
  arrange(Cruise, Station, lat, lon, MLD, par) %>% 
  mutate(p = zoo::na.approx(p, par, na.rm = F),
         ezd = ifelse(par == onePercSurfPAR, p, NA)) %>% 
  fill(ezd, .direction = "downup") %>% 
  select(-onePercSurfPAR) %>% 
  drop_na(s) %>% 
  ungroup() %>% 
  select(float, profile, datetime, ezd) %>% 
  distinct() %>% 
  left_join(floatsNmlds, .)  %>% 
  mutate( potT = swTheta(salinity = s, temperature = t, pressure = p),
          date = ymd(as_date(datetime)), 
          time = as_hms(datetime),
          plot_date = paste(month(date, label = T), day(date), format(parse_date_time(time, c('HMS', 'HM')), '%H:%M'))) %>% 
  select(float:datetime, date, time, plot_date, Cruise:t, potT, everything())


summary(floatsNmldsNezds$ezd)  
```


# T-S plots

We'll make T-S plots from the floats and overlay the data from the ship collected data. 

```{r}
ship_ctd <- read_rds("~/GITHUB/naames_multiday/Input/ctd_data.rds") %>% 
  filter(Cruise == "AT34" & Station == 4 | Cruise == "AT38" & Station %in% c(3,3.5,4,6)) %>% 
  select(CampCN:Station, `Longitude [degrees_east]`, lat, mld, pres_db, ave_temp_c, ave_sal_psu, fl_mg_m3) %>% 
  rename(datetime = Date, lon = `Longitude [degrees_east]`, p = pres_db, s = ave_sal_psu, t = ave_temp_c, MLD = mld, fchl = fl_mg_m3) %>% 
  mutate( potT = swTheta(salinity = s, temperature = t, pressure = p),
          date = ymd(as_date(datetime)), 
          time = as_hms(datetime),
          plot_date = paste(month(date, label = T), day(date), format(parse_date_time(time, c('HMS', 'HM')), '%H:%M'))) %>% 
  select(CampCN:datetime, date, time, plot_date, Cruise:t, potT, everything())
  
```

## Station 4

The float data for N2S4 shows different T-S at depth for May 29, so we'll exclude that date from further analysis. Also May 27 6:07 cast has different salinity at depth.


```{r message=FALSE, warning=FALSE}

s4_subset <- floatsNmldsNezds %>% filter(Cruise == "AT34", p <= 250,  datetime < ymd_hms("2016-05-31 00:00:00")) %>% mutate(type = "Float") %>% select(-c(o2:poc)) %>% distinct() %>% 
  bind_rows(., ship_ctd %>% filter(Cruise == "AT34", p <= 250) %>% mutate(type = "Ship")) %>% 
  filter(!plot_date %in% c("May 24 02:30", "May 27 06:07", "May 29 05:41", "May 29 20:41", "May 29 13:27"))

s4_subset$plot_date <- reorder(s4_subset$plot_date, s4_subset$datetime)

s4.ts <- PlotSvalbard::ts_plot(s4_subset, temp_col = "potT", sal_col = "s", color = "type", symbol_size = 3, symbol_shape = 16, symbol_alpha = 0.5, color_isopyc = "black", WM = NULL) +
  scale_color_viridis_d(begin = 0.1, end = 0.8) +
  theme_classic2(18) +
  scale_x_continuous(name = "Practical Salinity", sec.axis = sec_axis(~.,name = expression(paste("Potential Density, kg m"^-3)))) +
  scale_y_continuous(name = "Potential Temperature, ˚C") +
   theme(panel.spacing.x = unit(1, "cm"),
        axis.text.x = element_text(angle = 0),
        legend.title = element_blank(),
        legend.key.size = unit(0.4, "cm"),
        legend.position = c(0.8, 0.25),
        legend.text = element_text(size = 16),
        legend.background = element_rect(size = 0.2, linetype = "solid", color = "black"),
        legend.spacing.y = unit(0, "pt"))

s4_subset_ship <- s4_subset %>% filter(type == "Ship")

s4.ts2 <- PlotSvalbard::ts_plot(s4_subset_ship, temp_col = "potT", sal_col = "s", color = "plot_date", symbol_size = 3, symbol_shape = 16, symbol_alpha = 0.5, color_isopyc = "black", WM = NULL) +
  scale_color_viridis_d(begin = 0.1, end = 0.8) +
  theme_classic2(18) +
  scale_x_continuous(name = "Practical Salinity", sec.axis = sec_axis(~.,name = expression(paste("Potential Density, kg m"^-3)))) +
  scale_y_continuous(name = "Potential Temperature, ˚C") +
   theme(panel.spacing.x = unit(1, "cm"),
        axis.text.x = element_text(angle = 0),
        legend.title = element_blank(),
        legend.key.size = unit(0.4, "cm"),
        legend.position = c(0.8, 0.25),
        legend.text = element_text(size = 12),
        legend.background = element_rect(size = 0.2, linetype = "solid", color = "black"),
        legend.spacing.y = unit(0, "pt")) + 
  ggtitle("Ship")

s4_subset_float <- s4_subset %>% filter(type == "Float")

s4.ts3 <- PlotSvalbard::ts_plot(s4_subset_float, temp_col = "potT", sal_col = "s", color = "plot_date", symbol_size = 3, symbol_shape = 16, symbol_alpha = 0.5, color_isopyc = "black", WM = NULL) +
  scale_color_viridis_d(begin = 0.1, end = 0.8) +
  theme_classic2(18) +
  scale_x_continuous(name = "Practical Salinity", sec.axis = sec_axis(~.,name = expression(paste("Potential Density, kg m"^-3)))) +
  scale_y_continuous(name = "Potential Temperature, ˚C") +
   theme(panel.spacing.x = unit(1, "cm"),
        axis.text.x = element_text(angle = 0),
        legend.title = element_blank(),
        legend.key.size = unit(0.4, "cm"),
        legend.position = c(0.8, 0.25),
        legend.text = element_text(size = 12),
        legend.background = element_rect(size = 0.2, linetype = "solid", color = "black"),
        legend.spacing.y = unit(0, "pt")) +
  ggtitle("Float n0647")

```

```{r fig.height=10, fig.width=18}
s4.ts + s4.ts2 + s4.ts3
```

## Station 6

These casts have different T-S plots than all the float profiles and 7 of the ship profiles: plot_date %in% c("Sep 14 15:33", "Sep 15 03:04", "Sep 15 15:20", "Sep 16 03:04", "Sep 16 04:50", "Sep 16 07:26". We'll exclude these from further analysis. 

```{r message=FALSE, warning=FALSE}
s6_subset <- floatsNmldsNezds %>% filter(Cruise == "AT38" & Station == 6, p <= 250,  datetime < ymd_hms("2017-09-18 00:00:00")) %>%  mutate(type = "Float") %>% select(-c(o2:poc)) %>% distinct() %>% 
  bind_rows(., ship_ctd %>% filter(Cruise == "AT38", Station == 6, p <= 250) %>% mutate(type = "Ship")) %>% 
  filter(!plot_date %in% c("Sep 14 15:33", "Sep 15 03:04", "Sep 15 15:20", "Sep 16 03:04", "Sep 16 04:50", "Sep 16 07:26"))

s6_subset$plot_date <- reorder(s6_subset$plot_date, s6_subset$datetime)

s6.ts <- PlotSvalbard::ts_plot(s6_subset, temp_col = "potT", sal_col = "s", color = "type", symbol_size = 3, symbol_shape = 16, symbol_alpha = 0.7, color_isopyc = "black", WM = NULL) +
  scale_color_viridis_d(begin = 0.1, end = 0.8) +
  theme_classic2(18) +
  scale_x_continuous(name = "Practical Salinity", sec.axis = sec_axis(~.,name = expression(paste("Potential Density, kg m"^-3)))) +
  scale_y_continuous(name = "Potential Temperature, ˚C") +
   theme(panel.spacing.x = unit(1, "cm"),
        axis.text.x = element_text(angle = 0),
        legend.title = element_blank(),
        legend.key.size = unit(0.4, "cm"),
        legend.position = c(0.8, 0.7),
        legend.text = element_text(size = 12),
        legend.background = element_rect(size = 0.2, linetype = "solid", color = "black"),
        legend.spacing.y = unit(0, "pt"))

s6_subset_ship <- s6_subset %>% filter(type == "Ship")

s6.ts2 <- PlotSvalbard::ts_plot(s6_subset_ship, temp_col = "potT", sal_col = "s", color = "plot_date", symbol_size = 3, symbol_shape = 16, symbol_alpha = 0.7, color_isopyc = "black", WM = NULL) +
  scale_color_viridis_d(begin = 0.1, end = 0.8) +
  theme_classic2(18) +
  scale_x_continuous(name = "Practical Salinity", sec.axis = sec_axis(~.,name = expression(paste("Potential Density, kg m"^-3)))) +
  scale_y_continuous(name = "Potential Temperature, ˚C") +
   theme(panel.spacing.x = unit(1, "cm"),
        axis.text.x = element_text(angle = 0),
        legend.title = element_blank(),
        legend.key.size = unit(0.4, "cm"),
        legend.position = c(0.8, 0.7),
        legend.text = element_text(size = 12),
        legend.background = element_rect(size = 0.2, linetype = "solid", color = "black"),
        legend.spacing.y = unit(0, "pt")) +
  ggtitle("Ship")


s6_subset_float <- s6_subset %>% filter(type == "Float")

s6.ts3 <- PlotSvalbard::ts_plot(s6_subset_float, temp_col = "potT", sal_col = "s", color = "plot_date", symbol_size = 3, symbol_shape = 16, symbol_alpha = 0.7, color_isopyc = "black", WM = NULL) +
  scale_color_viridis_d(begin = 0.1, end = 0.8) +
  theme_classic2(18) +
  scale_x_continuous(name = "Practical Salinity", sec.axis = sec_axis(~.,name = expression(paste("Potential Density, kg m"^-3)))) +
  scale_y_continuous(name = "Potential Temperature, ˚C") +
   theme(panel.spacing.x = unit(1, "cm"),
        axis.text.x = element_text(angle = 0),
        legend.title = element_blank(),
        legend.key.size = unit(0.4, "cm"),
        legend.position = c(0.8, 0.7),
        legend.text = element_text(size = 12),
        legend.background = element_rect(size = 0.2, linetype = "solid", color = "black"),
        legend.spacing.y = unit(0, "pt")) + 
  ggtitle("Floats n0846 & n0847")
```

```{r fig.height=10, fig.width=18}
s6.ts + s6.ts2 + s6.ts3
```

## N3 Station 3 and 4

This T-S profile is different from the rest: filter(!plot_date %in% c("Sep 8 03:08")))

```{r message=FALSE, warning=FALSE}
s3_subset <- floatsNmldsNezds %>% filter(Cruise == "AT38" & Station == 3, p <= 250) %>%  mutate(type = "Float") %>% select(-c(o2:poc)) %>% distinct() %>% 
  bind_rows(., ship_ctd %>% filter(Cruise == "AT38", Station == 3, p <= 250) %>% mutate(type = "Ship") %>% filter(!plot_date %in% c("Sep 8 03:08")))

s3_subset$plot_date <- reorder(s3_subset$plot_date, s3_subset$datetime)

s3.ts <- PlotSvalbard::ts_plot(s3_subset, temp_col = "potT", sal_col = "s", color = "type", symbol_size = 3, symbol_shape = 16, symbol_alpha = 0.7, color_isopyc = "black", WM = NULL) +
  scale_color_viridis_d(begin = 0.1, end = 0.8) +
  theme_classic2(18) +
  scale_x_continuous(name = "Practical Salinity", sec.axis = sec_axis(~.,name = expression(paste("Potential Density, kg m"^-3)))) +
  scale_y_continuous(name = "Potential Temperature, ˚C") +
   theme(panel.spacing.x = unit(1, "cm"),
        axis.text.x = element_text(angle = 0),
        legend.title = element_blank(),
        legend.key.size = unit(0.4, "cm"),
        legend.position = c(0.8, 0.7),
        legend.text = element_text(size = 12),
        legend.background = element_rect(size = 0.2, linetype = "solid", color = "black"),
        legend.spacing.y = unit(0, "pt")) +
  ggtitle("NAAMES 3 Station 3 Float n0850")

n3s4_subset <- floatsNmldsNezds %>% filter(Cruise == "AT38" & Station == 4, p <= 250) %>%  mutate(type = "Float") %>% select(-c(o2:poc)) %>% distinct() %>% 
  bind_rows(., ship_ctd %>% filter(Cruise == "AT38", Station == 4, p <= 250) %>% mutate(type = "Ship") %>% filter(!plot_date %in% c("Sep 11 03:07")))

n3s4_subset$plot_date <- reorder(n3s4_subset$plot_date, n3s4_subset$datetime)

n3s4.ts <- PlotSvalbard::ts_plot(n3s4_subset, temp_col = "potT", sal_col = "s", color = "type", symbol_size = 3, symbol_shape = 16, symbol_alpha = 0.7, color_isopyc = "black", WM = NULL) +
  scale_color_viridis_d(begin = 0.1, end = 0.8) +
  theme_classic2(18) +
  scale_x_continuous(name = "Practical Salinity", sec.axis = sec_axis(~.,name = expression(paste("Potential Density, kg m"^-3)))) +
  scale_y_continuous(name = "Potential Temperature, ˚C") +
   theme(panel.spacing.x = unit(1, "cm"),
        axis.text.x = element_text(angle = 0),
        legend.title = element_blank(),
        legend.key.size = unit(0.4, "cm"),
        legend.position = c(0.8, 0.7),
        legend.text = element_text(size = 12),
        legend.background = element_rect(size = 0.2, linetype = "solid", color = "black"),
        legend.spacing.y = unit(0, "pt")) +
  ggtitle("NAAMES 3 Station 4 Float n0849")


s35_subset <- ship_ctd %>% filter(Cruise == "AT38", Station == 3.5, p <= 250) %>% mutate(type = "Ship")

s35_subset$plot_date <- reorder(s35_subset$plot_date, s35_subset$datetime)

s35.ts <- PlotSvalbard::ts_plot(s35_subset, temp_col = "potT", sal_col = "s", color = "plot_date", symbol_size = 3, symbol_shape = 16, symbol_alpha = 0.7, color_isopyc = "black", WM = NULL) +
  scale_color_viridis_d(begin = 0.1, end = 0.8) +
  theme_classic2(18) +
  scale_x_continuous(name = "Practical Salinity", sec.axis = sec_axis(~.,name = expression(paste("Potential Density, kg m"^-3)))) +
  scale_y_continuous(name = "Potential Temperature, ˚C") +
   theme(panel.spacing.x = unit(1, "cm"),
        axis.text.x = element_text(angle = 0),
        legend.title = element_blank(),
        legend.key.size = unit(0.4, "cm"),
        legend.position = c(0.8, 0.7),
        legend.text = element_text(size = 12),
        legend.background = element_rect(size = 0.2, linetype = "solid", color = "black"),
        legend.spacing.y = unit(0, "pt")) +
  ggtitle("NAAMES 3 Station 3.5")
```

```{r fig.height=10, fig.width=18}
s3.ts + s35.ts + n3s4.ts
```

# Map

```{r message=FALSE, warning=FALSE}
# chlaInfo <- info("erdMH1chla1day")
# n2_chl <- griddap(chlaInfo, latitude = c(46.5, 48.), longitude = c(-40.5, -37), time = c('2016-05-23','2016-05-27'), fields = 'chlorophyll')
# n3_chl <- griddap(chlaInfo, latitude = c(53.2, 53.5), longitude = c(-40, -39.4), time = c('2017-09-13', '2017-09-17'), fields = 'chlorophyll')


chlaInfo2 <- info("nesdisVHNSQchlaDaily")
n2_chl <- griddap(chlaInfo2, latitude = c(46.5, 48.), longitude = c(-40, -37), time = c('2016-05-23','2016-05-27'), fields = 'chlor_a')
n3_chl <- griddap(chlaInfo2, latitude = c(53.2, 53.5), longitude = c(-40, -39.4), time = c('2017-09-13', '2017-09-17'), fields = 'chlor_a')


sstInfo <- info("jplMURSST41")

n2_sst <- griddap(sstInfo, latitude = c(46.5, 48.), longitude = c(-40.5, -37), time = c('2016-05-24','2016-05-24'), fields = 'analysed_sst')
n3_sst <- griddap(sstInfo, latitude = c(53.2, 53.5), longitude = c(-40, -39.4), time = c('2017-09-13', '2017-09-13'), fields = 'analysed_sst')

```



```{r}
library(ggnewscale)

custom.shapes <- c("NAAMES 2 Station 4" = 21, "NAAMES 3 Station 3" = 22, "NAAMES 3 Station 3.5" = 23, "NAAMES 3 Station 4" = 24, "NAAMES 3 Station 6" = 25)
custom.colors <- c("Ship" = "#feb483", "Float n0847" = "#d31f2a", "Float n0846" = "#ffc000", "Float n0647" = "#27ab19", "Float n0849" = "#0db5e6", "Float n0850" = "#7139fe", "Drifter" = "#d16cfa")
custom.lines <- c("Ship" = 1, "Float n0847" = 2, "Float n0846" = 2, "Float n0647" = 2, "Float n0849" = 2 , "Float n0850" = 2, "Drifter" = 3)
mycolor <- colors$viridis
sst_color <- colors$temperature
w <- map_data("worldHires", ylim = c(35., 60.), xlim = c(-60, -30))


stations <- read_rds("~/GITHUB/naames_multiday/Input/ctd_data.rds") %>% 
  select(CampCN:lat) %>% 
  distinct() %>% 
  group_by(CampCN) %>% 
  filter(row_number() == 1) %>% 
  rename(datetime = `yyyy-mm-ddThh:mm:ss.sss`,
         lon = `Longitude [degrees_east]`) %>% 
  ungroup() 

map_data <- stations %>% select(Cruise, Station, Date, lon, lat) %>% mutate(type = "Ship") %>% 
  bind_rows(., floatsNmlds %>% select(Cruise, Station, float, datetime, lon, lat) %>% distinct() %>% rename(Date = datetime) %>%  mutate(type = paste("Float", float))) %>% 
  mutate(name = ifelse(Cruise == "AT34", "NAAMES 2", "NAAMES 3"), 
         plot_name = paste(name, "Station", Station),
         plot_sep = ifelse(Station == "6", "Subpolar", "Subtropical")) %>% 
  bind_rows(., drifters %>% mutate_at(vars(drifter), as.character) %>%  rename(float = drifter)) %>% 
  mutate(type = fct_relevel(type, 
            "Drifter", "Float n0647", "Float n0850", "Float n0849", "Float n0847", "Float n0846", "Ship"))

map1_data <- map_data %>% 
  filter(Cruise == "AT34")

map1 <- ggplot(data = map1_data, aes(x = lon, y = lat)) + 
  facet_wrap(~plot_sep, dir = "v", scales = "free") +
  # geom_tile(data = n2_sst$data, aes(x = lon, y = lat, fill = analysed_sst)) +
  geom_tile(data = n2_chl$data, aes(x = lon, y = lat, fill = chlor_a), alpha = 0.7) +
  scale_fill_gradientn(colors = mycolor, na.value = NA) +
  labs(fill = expression(paste("Chlorophyll", italic(" a"), ", mg m"^-3))) +
  new_scale_fill() +  
  geom_line(aes(color = type, group = interaction(type, plot_name, float), linetype = type), size = 1) +
  geom_point(aes(fill = type, shape = plot_name,  group = interaction(type, plot_name, float)),  size = 4, alpha = 0.8) +
  theme_linedraw() + 
  labs(x = "Longitude, ˚W", y = "Latitude, ˚N", color = "", fill = "", linetype = "", shape = "") +
  scale_linetype_manual(values = custom.lines) +
  scale_fill_manual(values = custom.colors) +
  scale_color_manual(values = custom.colors) +
  scale_shape_manual(values = custom.shapes) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  ggtitle("NAAMES 2 Station 4") 

map2_data <- map_data %>% 
  filter(Cruise == "AT38")

s4_map2 <- map1_data %>% 
  filter(type == "Ship")

map2 <- ggplot(data = map2_data, aes(x = lon, y = lat)) + 
  facet_wrap(~plot_sep, dir = "v", scales = "free") +
  # geom_tile(data = n3_sst$data, aes(x = lon, y = lat, fill = analysed_sst)) +
  geom_tile(data = n3_chl$data, aes(x = lon, y = lat, fill = chlor_a), alpha = 0.7) +
  scale_fill_gradientn(colors = mycolor, na.value = NA) +
  labs(fill = expression(paste("Chlorophyll", italic(" a"), ", mg m"^-3))) +
  new_scale_fill() +  
  geom_line(aes(color = type, group = interaction(type, plot_name, float), linetype = type), size = 1) +
  geom_point(aes(fill = type, shape = plot_name,  group = interaction(type, plot_name, float)),  size = 4, alpha = 0.8) +
  geom_line(data = s4_map2, aes(x = lon, y = lat, color = type, group = interaction(type, plot_name, float), linetype = type), size = 1) +
  geom_point(data = s4_map2, aes(x = lon, y = lat, fill = type, shape = plot_name), size = 4, alpha = 0.9) +
  theme_linedraw() + 
  labs(x = "Longitude, ˚W", y = "Latitude, ˚N", color = "", fill = "", linetype = "", shape = "") +
  scale_linetype_manual(values = custom.lines) +
  scale_fill_manual(values = custom.colors) +
  scale_color_manual(values = custom.colors) + 
  scale_shape_manual(values = custom.shapes) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  ggtitle("NAAMES 2 & 3")
```

```{r fig.height=6, fig.width=14}
(map1 + guides(fill = "none", color = "none", linetype = "none")) + (map2 + guides(fill = "none", shape = "none")) + plot_layout(guides = "collect")
```

#Station 4

## Temp profiles

```{r}
temp <- floatsNmldsNezds %>% 
  filter(!plot_date %in% c("May 29 05:41", "May 29 20:41", "May 29 13:27")) %>%  
   filter(Cruise == "AT34", p <= 300,  datetime < ymd_hms("2016-05-31 00:00:00")) %>%
  select(profile, datetime, p, t) %>% 
  bind_rows(., ship_ctd %>% 
              filter(!plot_date %in% c("May 24 02:30", "May 27 06:07")) %>% 
              filter(Cruise == "AT34" & Station == 4, p <= 300) %>% 
              select(datetime, p, t)) %>% 
  mutate(decimaldate = decimal_date(datetime)) %>% 
  select(decimaldate, p, t) 

mba <- MBA::mba.surf(temp, no.X = 300, no.Y = 300, extend = T)
dimnames(mba$xyz.est$z) <- list(mba$xyz.est$x, mba$xyz.est$y)
mba <- reshape2::melt(mba$xyz.est$z, varnames = c('decimal_date', 'p'), value.name = 'temp')
```

### Plot

```{r fig.align="center", fig.height=10, fig.width=16, echo = FALSE, message=FALSE, warning=FALSE}

t_odv_data <- mba %>% 
  mutate(datetime = date_decimal(decimal_date)) 

profile_points <- floatsNmldsNezds %>% 
  filter(!plot_date %in% c("May 29 05:41", "May 29 20:41", "May 29 13:27")) %>% 
  filter(Cruise == "AT34", p <= 300,  datetime < ymd_hms("2016-05-31 00:00:00")) %>%
  select(float, profile, datetime, MLD, p, t, s, fchl, ezd) %>% 
  mutate(type = "Float") %>% 
  bind_rows(., ship_ctd %>% 
               filter(!plot_date %in% c("May 24 02:30", "May 27 06:07")) %>% 
              filter(Cruise == "AT34" & Station == 4, p <= 300) %>% 
              select(CampCN, datetime, MLD, p, t, s, fchl) %>% 
              mutate(type = "Ship")) 
  
t.plot <- t_odv_data %>% 
  ggplot(aes(x = datetime, y = p)) + 
  geom_raster(aes(fill = temp)) +
  geom_line(data = profile_points, aes(x = datetime, y = p, linetype = type, group = interaction(type, datetime)), alpha = 0.4, size = 1) +
  geom_line(data = profile_points, aes(x = datetime, y = MLD), color = "White", linetype = 1, size = 1.5 ) +
  geom_line(data = profile_points[!is.na(profile_points$ezd),], aes(x = datetime, y = ezd), color = "Red", size = 1.5 ) +
  scale_fill_gradientn(colors = sst_color, na.value = NA) +
  scale_y_reverse() +
  geom_contour(aes(z = temp), binwidth = 0.1, colour = "black", alpha = 0.2) +
  labs(y = "Depth, m", x = "", fill = expression(paste("Temperature, ˚C")), shape = "", linetype = "") +
  guides(fill = guide_colourbar(direction = "horizontal", barheight = 2, barwidth = 20, order = 2)) +
  theme_classic2(18) +
  theme(legend.position = "top") 

```


## Sal profiles

```{r}
sal <- floatsNmldsNezds %>% 
  filter(!plot_date %in% c("May 29 05:41", "May 29 20:41", "May 29 13:27")) %>%  
   filter(Cruise == "AT34", p <= 300,  datetime < ymd_hms("2016-05-31 00:00:00")) %>%
  select(profile, datetime, p, s) %>% 
  bind_rows(., ship_ctd %>% 
              filter(!plot_date %in% c("May 24 02:30", "May 27 06:07")) %>% 
              filter(Cruise == "AT34" & Station == 4, p <= 300) %>% 
              select(datetime, p, s)) %>% 
  mutate(decimaldate = decimal_date(datetime)) %>% 
  select(decimaldate, p, s) 

mba <- MBA::mba.surf(sal, no.X = 300, no.Y = 300, extend = T)
dimnames(mba$xyz.est$z) <- list(mba$xyz.est$x, mba$xyz.est$y)
mba <- reshape2::melt(mba$xyz.est$z, varnames = c('decimal_date', 'p'), value.name = 's')
```

### Plot

```{r fig.align="center", fig.height=10, fig.width=16, echo = FALSE, message=FALSE, warning=FALSE}
sal_colors <- colors$salinity

s_odv_data <- mba %>% 
  mutate(datetime = date_decimal(decimal_date)) 
  
s.plot <- s_odv_data %>% 
  ggplot(aes(x = datetime, y = p)) + 
  geom_raster(aes(fill = s)) +
  geom_line(data = profile_points, aes(x = datetime, y = p, linetype = type, group = interaction(type, datetime)), alpha = 0.4, size = 1) +
  geom_line(data = profile_points, aes(x = datetime, y = MLD), color = "White", linetype = 1, size = 1.5 ) +
  geom_line(data = profile_points[!is.na(profile_points$ezd),], aes(x = datetime, y = ezd), color = "Red", size = 1.5 ) +
  scale_fill_gradientn(colors = sal_colors, na.value = NA) +
  scale_y_reverse() +
  geom_contour(aes(z = s), binwidth = 0.05, colour = "black", alpha = 0.2) +
  labs(y = "Depth, m", x = "", fill = expression(paste("Salinity, ppt")), shape = "", linetype = "") +
  guides(fill = guide_colourbar(direction = "horizontal", barheight = 2, barwidth = 20, order = 2)) +
  theme_classic2(18) +
  theme(legend.position = "top") 

```


##  Fluorescence profiles

```{r}
fl <- floatsNmldsNezds %>% 
  filter(!plot_date %in% c("May 29 05:41", "May 29 20:41", "May 29 13:27")) %>%  
  filter(Cruise == "AT34", p <= 300,  datetime < ymd_hms("2016-05-31 00:00:00")) %>%
  select(profile, datetime, p, fchl) %>% 
  bind_rows(., ship_ctd %>% 
              filter(!plot_date %in% c("May 24 02:30","May 27 06:07")) %>% 
              filter(Cruise == "AT34" & Station == 4, p <= 300) %>% 
              select(datetime, p, fchl)) %>% 
  mutate(decimaldate = decimal_date(datetime)) %>% 
  select(decimaldate, p, fchl) 

mba <- MBA::mba.surf(fl, no.X = 300, no.Y = 300, extend = T)
dimnames(mba$xyz.est$z) <- list(mba$xyz.est$x, mba$xyz.est$y)
mba <- reshape2::melt(mba$xyz.est$z, varnames = c('decimal_date', 'p'), value.name = 'fchl')
```

### Plot

```{r fig.align="center", fig.height=10, fig.width=16, echo = FALSE, message=FALSE, warning=FALSE}

fl_odv_data <- mba %>% 
  mutate(datetime = date_decimal(decimal_date)) 
  
fl.plot <- fl_odv_data %>% 
  ggplot(aes(x = datetime, y = p)) + 
  geom_raster(aes(fill = fchl)) +
  geom_line(data = profile_points, aes(x = datetime, y = p, linetype = type, group = interaction(type, datetime)), alpha = 0.4, size = 1) +
 geom_line(data = profile_points, aes(x = datetime, y = MLD), color = "White", linetype = 1, size = 1.5 ) +
  geom_line(data = profile_points[!is.na(profile_points$ezd),], aes(x = datetime, y = ezd), color = "Red", size = 1.5 ) +
  scale_fill_gradientn(colors = mycolor, na.value = NA) +
  scale_y_reverse() +
  geom_contour(aes(z = fchl), binwidth = 0.5, colour = "black", alpha = 0.2) +
  labs(y = "Depth, m", x = "", fill = expression(paste("Chl", italic(" a"), " Fluorescence, mg m"^-3)), shape = "", linetype = "") +
  guides(fill = guide_colourbar(direction = "horizontal", barheight = 2, barwidth = 20, order = 2)) +
  theme_classic2(18) +
  theme(legend.position = "top")

```

```{r fig.height=16, fig.width=12}
t.plot / s.plot / fl.plot
```


# Station 6

These plots represent data from two separate floats. The T-S plots shoow good agreement between the water masses sampled by the floats. 

## Temp profiles

```{r}
temp <- floatsNmldsNezds %>% 
   filter(Cruise == "AT38", Station == 6, p <= 300,  datetime < ymd_hms("2017-09-18 00:00:00")) %>%
  select(profile, datetime, p, t) %>% 
  bind_rows(., ship_ctd %>% 
               filter(!plot_date %in% c("Sep 14 15:33", "Sep 15 03:04", "Sep 15 15:20", "Sep 16 03:04", "Sep 16 04:50", "Sep 16 07:26")) %>% 
              filter(Cruise == "AT38" & Station == 6, p <= 300) %>% 
              select(datetime, p, t)) %>% 
  mutate(decimaldate = decimal_date(datetime)) %>% 
  select(decimaldate, p, t) 

mba <- MBA::mba.surf(temp, no.X = 300, no.Y = 300, extend = T)
dimnames(mba$xyz.est$z) <- list(mba$xyz.est$x, mba$xyz.est$y)
mba <- reshape2::melt(mba$xyz.est$z, varnames = c('decimal_date', 'p'), value.name = 'temp')
```

### Plot

```{r fig.align="center", fig.height=10, fig.width=16, echo = FALSE, message=FALSE, warning=FALSE}

t_odv_data <- mba %>% 
  mutate(datetime = date_decimal(decimal_date)) 

profile_points <- floatsNmldsNezds %>% 
  filter(Cruise == "AT38", Station == 6, p <= 300,  datetime < ymd_hms("2017-09-18 00:00:00")) %>%
  select(float, profile, datetime, MLD, ezd, p, t, s, fchl) %>% 
  mutate(type = "Float") %>% 
  bind_rows(., ship_ctd %>% 
              filter(!plot_date %in% c("Sep 14 15:33", "Sep 15 03:04", "Sep 15 15:20", "Sep 16 03:04", "Sep 16 04:50", "Sep 16 07:26")) %>% 
              filter(Cruise == "AT38" & Station == 6, p <= 300) %>% 
              select(CampCN, datetime, MLD, p, t, s, fchl) %>% 
              mutate(type = "Ship")) 
  
t.plot <- t_odv_data %>% 
  ggplot(aes(x = datetime, y = p)) + 
  geom_raster(aes(fill = temp)) +
  geom_line(data = profile_points, aes(x = datetime, y = p, linetype = type, group = interaction(type, datetime)), alpha = 0.4, size = 1) +
  geom_line(data = profile_points, aes(x = datetime, y = MLD), color = "White", linetype = 1, size = 1.5 ) +
  geom_line(data = profile_points[!is.na(profile_points$ezd),], aes(x = datetime, y = ezd), color = "Red", size = 1.5 ) +
  scale_fill_gradientn(colors = sst_color, na.value = NA) +
  scale_y_reverse() +
  geom_contour(aes(z = temp), binwidth = 0.1, colour = "black", alpha = 0.2) +
  labs(y = "Depth, m", x = "", fill = expression(paste("Temperature, ˚C")), shape = "", linetype = "") +
  guides(fill = guide_colourbar(direction = "horizontal", barheight = 2, barwidth = 20, order = 2)) +
  theme_classic2(18) +
  theme(legend.position = "top") 

```


## Sal profiles

```{r}
sal <- floatsNmldsNezds %>% 
   filter(Cruise == "AT38", Station == 6, p <= 300,  datetime < ymd_hms("2017-09-18 00:00:00")) %>%
  select(profile, datetime, p, s) %>% 
  bind_rows(., ship_ctd %>% 
              filter(!plot_date %in% c("Sep 14 15:33", "Sep 15 03:04", "Sep 15 15:20", "Sep 16 03:04", "Sep 16 04:50", "Sep 16 07:26")) %>% 
              filter(Cruise == "AT38" & Station == 6, p <= 300) %>% 
              select(datetime, p, s)) %>% 
  mutate(decimaldate = decimal_date(datetime)) %>% 
  select(decimaldate, p, s) 

mba <- MBA::mba.surf(sal, no.X = 300, no.Y = 300, extend = T)
dimnames(mba$xyz.est$z) <- list(mba$xyz.est$x, mba$xyz.est$y)
mba <- reshape2::melt(mba$xyz.est$z, varnames = c('decimal_date', 'p'), value.name = 's')
```

### Plot

```{r fig.align="center", fig.height=10, fig.width=16, echo = FALSE, message=FALSE, warning=FALSE}
sal_colors <- colors$salinity

s_odv_data <- mba %>% 
  mutate(datetime = date_decimal(decimal_date)) 
  
s.plot <- s_odv_data %>% 
  ggplot(aes(x = datetime, y = p)) + 
  geom_raster(aes(fill = s)) +
  geom_line(data = profile_points, aes(x = datetime, y = p, linetype = type, group = interaction(type, datetime)), alpha = 0.4, size = 1) +
   geom_line(data = profile_points, aes(x = datetime, y = MLD), color = "White", linetype = 1, size = 1.5 ) +
  geom_line(data = profile_points[!is.na(profile_points$ezd),], aes(x = datetime, y = ezd), color = "Red", size = 1.5 ) +
  scale_fill_gradientn(colors = sal_colors, na.value = NA) +
  scale_y_reverse() +
  geom_contour(aes(z = s), binwidth = 0.05, colour = "black", alpha = 0.2) +
  labs(y = "Depth, m", x = "", fill = expression(paste("Salinity, ppt")), shape = "", linetype = "") +
  guides(fill = guide_colourbar(direction = "horizontal", barheight = 2, barwidth = 20, order = 2)) +
  theme_classic2(18) +
  theme(legend.position = "top") 

```




##  Fluorescence profiles

```{r}
fl <- floatsNmldsNezds %>% 
  filter(Cruise == "AT38", Station == 6, p <= 300,  datetime < ymd_hms("2017-09-18 00:00:00")) %>%
  select(profile, datetime, p, fchl) %>% 
  bind_rows(., ship_ctd %>% 
              filter(!plot_date %in% c("Sep 14 15:33", "Sep 15 03:04", "Sep 15 15:20", "Sep 16 03:04", "Sep 16 04:50", "Sep 16 07:26")) %>% 
              filter(Cruise == "AT38" & Station == 6, p <= 300) %>% 
              select(datetime, p, fchl)) %>% 
  mutate(decimaldate = decimal_date(datetime)) %>% 
  select(decimaldate, p, fchl) 

mba <- MBA::mba.surf(fl, no.X = 300, no.Y = 300, extend = T)
dimnames(mba$xyz.est$z) <- list(mba$xyz.est$x, mba$xyz.est$y)
mba <- reshape2::melt(mba$xyz.est$z, varnames = c('decimal_date', 'p'), value.name = 'fchl')
```

### Plot

```{r fig.align="center", fig.height=10, fig.width=16, echo = FALSE, message=FALSE, warning=FALSE}

fl_odv_data <- mba %>% 
  mutate(datetime = date_decimal(decimal_date)) 
  
fl.plot <- fl_odv_data %>% 
  ggplot(aes(x = datetime, y = p)) + 
  geom_raster(aes(fill = fchl)) +
  geom_line(data = profile_points, aes(x = datetime, y = p, linetype = type, group = interaction(type, datetime)), alpha = 0.4, size = 1) +
   geom_line(data = profile_points, aes(x = datetime, y = MLD), color = "White", linetype = 1, size = 1.5 ) +
  geom_line(data = profile_points[!is.na(profile_points$ezd),], aes(x = datetime, y = ezd), color = "Red", size = 1.5 ) +
  scale_fill_gradientn(colors = mycolor, na.value = NA) +
  scale_y_reverse() +
  geom_contour(aes(z = fchl), binwidth = 0.5, colour = "black", alpha = 0.2) +
  labs(y = "Depth, m", x = "", fill = expression(paste("Chl", italic(" a"), " Fluorescence, mg m"^-3)), shape = "", linetype = "") +
  guides(fill = guide_colourbar(direction = "horizontal", barheight = 2, barwidth = 20, order = 2)) +
  theme_classic2(18) +
  theme(legend.position = "top")

```

```{r fig.height=16, fig.width=12}
t.plot / s.plot / fl.plot
```

# Save data

```{r}
write_rds(floatsNmldsNezds, "~/GITHUB/naames_multiday/Output/processed_floats.rds")
```

